---
title: "Introduction to miter"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to miter}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Overview

The **miter** package provides tools for iterative modeling and forecasting with tidymodels. It facilitates creating nested tables for grouped time series data, adding multiple modeling workflows, performing cross-validation, fitting models, generating predictions, and creating ensemble models.

## Installation

```{r, eval=FALSE}
# Install from GitHub
devtools::install_github("regisely/miter")
```

## Load Package

```{r setup, message=FALSE}
library(miter)
library(dplyr)
library(ggplot2)
```

# Getting Started

## Example Data

The package includes `icms_br`, a dataset of Brazilian state tax revenue:

```{r, eval=FALSE}
data(icms_br)
head(icms_br)

# View structure
str(icms_br)

# Check date range
range(icms_br$date)

# Number of states
length(unique(icms_br$uf))
```

## Basic Workflow

### 1. Create a Miter Table

A miter table is a nested tibble that groups data by specified variables:

```{r, eval=FALSE}
tbl <- icms_br %>%
  filter(uf %in% c("SP", "RJ")) %>%
  group_by(uf) %>%
  miter_table()

tbl
```

### 2. Initialize and Add Workflows

Create modeling workflows for your time series:

```{r, eval=FALSE}
# Initialize time series models
workflows <- initialize_ts_models(
  data = icms_br,
  outcome_var = "value",
  id_var = "uf",
  date_var = "date"
)

# Add workflows to table
tbl <- tbl %>%
  add_workflows(workflows)

tbl
```

### 3. Create Data Splits

Create holdout splits for train/test evaluation:

```{r, eval=FALSE}
tbl <- tbl %>%
  holdout_time_split(prop = 0.8)

tbl
```

### 4. Fit Models

Fit all workflows to your data:

```{r, eval=FALSE}
tbl <- tbl %>%
  fit(splits, control = control_miter(progress = TRUE))

tbl
```

### 5. Generate Predictions

Extract predictions from fitted models:

```{r, eval=FALSE}
preds <- tbl %>%
  predict(fitted_splits)

head(preds)
```

### 6. Calculate Metrics

Evaluate model performance:

```{r, eval=FALSE}
metrics <- calculate_metrics(preds)
metrics
```

### 7. Visualize Results

Plot predictions and metrics:

```{r, eval=FALSE}
# Plot predictions
autoplot(preds)

# Plot metrics
autoplot(metrics)
```

# Advanced Features

## Cross-Validation

Create resamples for hyperparameter tuning:

```{r, eval=FALSE}
tbl <- tbl %>%
  cv_time_split(
    column = splits,
    initial = 36,
    assess = 12,
    slice_limit = 10
  ) %>%
  fit(resamples_train)
```

## Nested Cross-Validation

For more robust evaluation:

```{r, eval=FALSE}
tbl <- tbl %>%
  nested_cv_time_split(data) %>%
  fit(nested_cv_data)
```

## Ensemble Models

Create ensemble models from base model predictions:

```{r, eval=FALSE}
# Initialize ensemble workflows
ensemble_workflows <- initialize_ensemble_models(preds, "date")

# Create ensemble (using top 5 models)
ensemble <- create_ensemble(
  data = preds,
  workflows = ensemble_workflows,
  n_models = 5
)

# Add ensemble predictions
combined_preds <- add_ensemble(ensemble, preds)

# Evaluate
metrics_combined <- calculate_metrics(combined_preds)
autoplot(metrics_combined)
```

## Comprehensive Model Set

Use a wider range of models:

```{r, eval=FALSE}
workflows_all <- initialize_all_models(
  data = icms_br,
  outcome_var = "value",
  id_var = "uf",
  date_var = "date",
  outcome_lags = 3
)

# Use ARIMA specifications fitted to data
workflows_fixed <- initialize_all_models(
  data = icms_br,
  outcome_var = "value",
  id_var = "uf",
  date_var = "date",
  outcome_lags = 3,
  fix_arima_specs = TRUE
)
```

## Custom Control Parameters

Fine-tune the fitting process:

```{r, eval=FALSE}
# Standard control
ctrl <- control_miter(
  progress = TRUE,
  verbose = FALSE,
  save_pred = TRUE,
  allow_par = TRUE
)

# Racing control (with early stopping)
ctrl_race <- control_miter_race(
  burn_in = 5,
  alpha = 0.01,
  num_ties = 10
)

# Use in fitting
tbl <- tbl %>%
  fit(resamples_train, control = ctrl_race)
```

# Tips and Best Practices

1. **Start Small**: Begin with a subset of your data and a few models to test your workflow

2. **Use Progress Bars**: Set `progress = TRUE` in `control_miter()` to monitor long-running fits

3. **Save Intermediate Results**: Store fitted tables and predictions as you go

4. **Monitor Memory**: With many models and large datasets, consider processing in batches

5. **Leverage Parallelization**: Set `allow_par = TRUE` and configure a parallel backend for faster fitting

6. **Validate on Holdout**: Always create a final holdout set that isn't used during development

7. **Compare Metrics**: Use `calculate_metrics()` with different metric sets to get a complete picture

# Conclusion

The miter package streamlines the process of fitting and evaluating multiple models across grouped time series data. By leveraging the tidymodels ecosystem, it provides a consistent and efficient interface for iterative forecasting workflows.

For more information, see the package documentation with `?miter` or visit the GitHub repository.
